---
title: "Investigating America's Most Reputable Companies"
subtitle: "STA/ISS 313 - Spring 2023 - Project 1"
author: "Team Six: George McCabe, Shelby Tisdale, Florence Wang, Chris Yang"
title-slide-attributes:
  data-background-size: stretch
  data-slide-number: none
format: revealjs
editor: visual
execute:
  echo: false
---

```{r}
#| label: load-packages
#| include: false

# Add any additional packages you need to this chunk
# Remove any packages from this list that you're not using
library(tidyverse)
library(tidymodels)
library(gt)
library(patchwork)
library(gganimate)
library(ggcorrplot)
library(dplyr)
library(ggplot2)
library(patchwork)
```

```{r}
#| label: setup
#| include: false

# For better figure resolution
knitr::opts_chunk$set(
  fig.retina = 3, 
  dpi = 300, 
  fig.width = 6, 
  fig.asp = 0.618, 
  out.width = "70%"
  )
```

```{r}
#| label: load-data
#| include: false

poll <- read.csv("data/polls.csv")
reputation <- read.csv("data/reputation.csv")
```

# Investigating America's Most Reputable Companies

## Introduction to Dataset and Motivation

-   The presentation is created using the Quarto CLI

-   `##` sets the start of a new slide

## Question 1: How do customer ratings of company attributes vary by industry and how do these attributes relate to one another?

### Visualization #1

```{r}
#| label: data wrangling q1v1
#| echo: false
#data wrangling
reputation1 <- reputation |>
  group_by(industry, name) |>
  mutate(avg_score = mean(score))
reputation2 <- reputation1 |>
  group_by(name) |>
  mutate(maxavg = max(avg_score))

#used to confirm which industries were the highest scored in each category
overview <- reputation1 |>
  group_by(name) |>
  slice(which.max(avg_score))

#creating dataframes for each attribute based on which industry had the highest average
citizenship <- reputation2 |>
  group_by(name) |>
  filter(name == "CITIZENSHIP" & avg_score == maxavg)
culture <- reputation2 |>
  group_by(name) |>
  filter(name == "CULTURE" & avg_score == maxavg)
ethics <- reputation2 |>
  group_by(name) |>
  filter(name == "ETHICS" & avg_score == maxavg)
growth <- reputation2 |>
  group_by(name) |>
  filter(name == "GROWTH" & avg_score == maxavg)
ps <- reputation2 |>
  group_by(name) |>
  filter(name == "P&S" & avg_score == maxavg)
trust <- reputation2 |>
  group_by(name) |>
  filter(name == "TRUST" & avg_score == maxavg)
vision <- reputation2 |>
  group_by(name) |>
  filter(name == "VISION" & avg_score == maxavg)

final <- rbind(citizenship, culture, ethics, growth, ps, trust, vision)
```

```{r}
#| label: q1v1 plot
#| echo: false
get_box_stats <- function(y, upper_limit = max(final$score) * 1.15) {
  return(data.frame(
    y = 0.95 * upper_limit,
    label = paste(
      "Mean =", round(mean(y), 2), "\n",
      "Med =", round(median(y), 2), "\n"
    )
  ))
}
ggplot(final, aes(x = name, y = score, fill = industry)) +
  geom_boxplot(fatten = 0.8) +
  labs(title = "Highest Rated Industries In Each Attribute Category", x = "Company Attribute", y = "Score", fill = "Industry") +
  scale_fill_manual(values = c("#4ea5e3", "#ce0e5b", "#1aaf82")) +
  stat_summary(fun.data = get_box_stats, geom = "text", hjust = 0.5, vjust = 0.9, size = 2.6) +
  theme_minimal() 
```

## Question 1 Continued

### Visualization #2

```{r}
#| label: data wrangling q1 v2
#| echo: false

reputation_longer <- reputation |>
  select(company, name, score, industry) |>
  pivot_wider(
    names_from = name,
    values_from = score
  )

reputation_longer <- reputation_longer |>
  rename("Trust" = "TRUST",
         "Ethics" = "ETHICS",
         "Growth" = "GROWTH",
         "Citizenship" = "CITIZENSHIP",
         "Vision" = "VISION",
         "Culture" = "CULTURE")

attributes_corr <- reputation_longer |>
  as.data.frame() |>               
  select(Trust, Ethics, Growth, `P&S`, Citizenship, Vision, Culture) |>                   
  as.matrix() |>                   
  cor() 
attributes_p_values <- cor_pmat(attributes_corr)
```

```{r}
#| label: q1v2 plot
#| echo: false
#| warning: false
#| fig-alt: |
#|   Body mass vs. flipper lenght of Palmer Penguins for species Adelie and all 
#|   the others combined together. There is a relatively strong, positive 
#|   relationship between the two variables. The Adelie penguins are clustered 
#|   together but they don't exhibit a different trend than the rest of the 
#|   penguins.

ggcorrplot(attributes_corr,
           type = "lower",
           method = "circle",
           show.legend = TRUE,
           legend.title = "Correlation Coefficent",
           p.mat = attributes_p_values,
           colors = c("#ce0e5b", "white", "#4ea5e3")) + 
  annotate("text",
           x = 2.7, y = 6, 
           color = "#b320ab", size = 3.8,
           label = "Size of circle indicates 
strength of positive correlation") +
    annotate("text",
           x = 2.7, y = 5, 
           color = "#1aaf82", size = 3.8,
           label = "X's indicate no significant 
relationship (alpha > 0.05)") +
  labs(
    title = "Correlation Matrix of Company Attributes",
    subtitle = "Attribute scores from 2022"
  ) +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))
```

## Question 2: How does company RQ score change between 2017-2022 and are the trends consistent with the overall industry?

### Visualization #1

```{r}
#| label: q2v1 plot

# filter NAs
poll <- poll |>
  group_by(company) |>
  filter(!any(is.na(rq) & year %in% 2017:2021)) |>
  ungroup()

# remove unnecessary column
poll <- select(poll, -'2022_rank',  -'change', -'rank')

# create year 2022 column with 2022 RQ score 
poll_2022 <- poll |>
  select(company, '2022_rq', 'industry') |>
  distinct() |>
  rename(rq = `2022_rq`) |>
  mutate(year = 2022)

# merge two datasets together
poll <- bind_rows(poll, poll_2022)

# remove unnecessary column
poll <- select(poll, -'2022_rq')

# calculate RQ score change from 2017-2022 
poll_change <- poll |> 
  group_by(company) |> 
  summarize(rq_change = rq[year == 2022] - rq[year == 2017])

# Get the top 3 companies with the greatest increase and decrease in RQ score
poll_top3_inc <- poll_change |> 
  arrange(desc(rq_change)) |> 
  slice(1:3)

poll_top3_dec <- poll_change |> 
  arrange(rq_change) |> 
  slice(1:3)

# select top 3 companies in each 
companies <- poll |>
  filter(company %in% c("Johnson & Johnson", "Goldman Sachs", "Facebook", "Samsung", "Wells Fargo & Company", "The Walt Disney Company"))
```

```{r}

# plot animation graph 
company_plot <- ggplot(companies, aes(x = year, y = rq, group = company, color = industry)) +
  geom_line(size = 1) +
  geom_point(size = 3) +
  geom_text(aes(label = company), hjust = -0.1, vjust = -0.5, color = "black", size = 2) +
  theme_bw() +
  labs(title = "Top Companies with Largest Increase and Decrease in RQ Score",
       subtitle = "from 2017-2022", 
       x = "Year", 
       y = "RQ Score") + 
  scale_color_manual(values=c(c("#F9C80E", "#EA3546", "#662E9B", "#43BCCD"))) +
  transition_reveal(year) +
  ease_aes('linear') + 
  theme_bw()

animation <- animate(company_plot, fps = 1)
animation
```

## Question 2 Continued

### Visualization #2

```{r}
#| label: question 2 vizualization 2
#| message: false
#| warning: false

# data set with company, rq, and year for years before 2022
before22_poll <- poll |>
  select(company, rq, year, industry) |>
  drop_na(rq)

# data set with company, rq, and year for 2022
`2022_poll` <- poll |>
  select(company, `X2022_rq`) |>
  drop_na(`X2022_rq`) |>
  pivot_longer(
    cols = `X2022_rq`,
    names_to = "year",
    values_to = "rq"
  ) |>
  mutate(year = parse_number(year))

# combine into new dataset
new_poll <- full_join(before22_poll, `2022_poll`, by = c("company", "rq", "year"))

# rq in 2017
`17_rq` <- new_poll |>
  filter(year == 2017) |>
  select(company, rq) |>
  rename("start_rq" = "rq")

# rq in 2022
`22_rq` <- new_poll |>
  filter(year == 2022) |>
  select(company, rq) |>
  rename("end_rq" = "rq")

# change in rq
rq_change <- new_poll |>
  select(company, industry) |>
  distinct(company) |>
  right_join(`17_rq`, by = "company") |>
  right_join(`22_rq`, by = "company") |>
  mutate(rq_change = end_rq - start_rq) |>
  distinct(company, .keep_all = TRUE) |> 
  select(company, rq_change) |>
  drop_na(rq_change)

# find 3 companies greatest rq increase
top3_rq <- rq_change |>
  arrange(rq_change) |>
  slice_head(n = 3)

# find 3 companies greatest rq decrease
last3_rq <- rq_change |>
  arrange(rq_change) |>
  slice_tail(n = 3)

# companies
companies <- c("Samsung", "Goldman Sachs", "Wells Fargo & Company",
               "Facebook", "Johnson & Johnson", "The Walt Disney Company")

# industries that these companies are in 
industries <- c("Tech", "Media", "Pharma", "Financial Services")

# data set for geom line
line_data <- new_poll |> filter(company %in% companies) |>
    mutate(industry = case_when(company == "Samsung" ~ "Tech",
                                company == "The Walt Disney Company" ~ "Media",
                                company == "Johnson & Johnson" ~  "Pharma",
                                company == "Goldman Sachs" ~ "Financial Services",
                                company == "Wells Fargo & Company" ~ "Financial Services",
                                company == "Facebook" ~  "Tech")) |>
    group_by(industry, year) |>
    summarize(mean_rq = mean(rq), .groups = "drop")

# fill in mising industry values for 2022 data
new_poll <- new_poll |>
  mutate(industry = case_when(
    company == "Apple" ~ "Tech",
    company == "American Express" ~ "Financial Services",
    company == "Bank of America" ~ "Financial Services",
    company == "Citigroup" ~ "Financial Services",
    company == "Comcast" ~ "Media",
    company == "Dell" ~ "Tech",
    company == "Electronic Arts, Inc." ~ "Tech",
    company == "Facebook" ~ "Tech",
    company == "Fox Corporation" ~ "Media",
    company == "Goldman Sachs" ~ "Financial Services",
    company == "Google" ~ "Tech",
    company == "IBM" ~ "Tech",
    company == "Johnson & Johnson" ~ "Pharma",
    company == "JPMorgan Chase & Co." ~ "Financial Services",
    company == "LG Corporation" ~ "Tech",
    company == "Microsoft" ~ "Tech",
    company == "Netflix" ~ "Tech",
    company == "Pfizer" ~ "Pharma",
    company == "Robinhood" ~ "Tech",
    company == "Samsung" ~ "Tech",
    company == "Sony" ~ "Tech",
    company == "Spotify" ~ "Media",
    company == "State Farm Insurance" ~ "Financial Services",
    company == "The Walt Disney Company" ~ "Media",
    company == "TikTok" ~ "Tech",
    company == "Twitter" ~ "Tech",
    company == "Uber" ~ "Tech",
    company == "USAA" ~ "Financial Services",
    company == "Wells Fargo & Company" ~ "Financial Services",
    TRUE ~ industry
  ))


# create visualization
new_poll |> 
  filter(industry %in% industries) |>
  ggplot(aes(x = year, y = rq, group = industry, color = industry)) +
  geom_smooth(method = 'loess',
              alpha = 0.25,
              linewidth = 0.5,
              aes(fill = industry, color = industry, linetype = "company fit")) +
  geom_line(data = line_data, show.legend = FALSE,
            aes(x = year, y = mean_rq, group = industry, color = industry,
                linetype = "industry mean"),
            linewidth = 1) +
  facet_wrap(~ industry, ncol = 4) +
  geom_point(data = line_data,
             aes(x = year, y = mean_rq), size = 3) +
  theme_minimal() +
  labs(title = "Mean Company RQ Changes Over Time for Different Industries",
       y = "RQ",
       x = "Year",
       color = "Industry",
       fill = "Industry",
       subtitle = "with company trendlines and variability in RQ scores",
       linetype = "Line") +
  scale_color_viridis_d() + 
  scale_color_manual(values = c("#F9C80E", "#EA3546", "#662E9B", "#43BCCD")) +
  scale_fill_manual(values = c("#F9C80E", "#EA3546", "#662E9B", "#43BCCD") )+
  theme(legend.position = "bottom",
        axis.text = element_text(size = 7),
        panel.spacing=unit(1,"lines"),
        legend.direction = "vertical",
        legend.box = "horizontal"
        ) +
  scale_linetype_manual(values = c("dotted", "solid"))

```
