---
title: "Project title"
subtitle: "Proposal"
format: html
editor: visual
---

```{r}
#| label: load-pkgs
#| message: false

library(tidyverse)
```

# Dataset

A brief description of your dataset including its provenance, dimensions, etc. as well as the reason why you chose this dataset.

Make sure to load the data and use inline code for some of this information.

```{r loading poll data}
write_csv(poll, "/home/guest/project-01-company_reputation_poll/polls.csv")
```

```{r loading reputation data}
write_csv(reputation, "/home/guest/project-01-company_reputation_poll/reputation.csv")
```

Our dataset comes we are all consumers from many of these industries company reputations blah blah

# Questions

1.  Which industry on average ranked the highest for each rating attribute in the year of 2022?

2.  How does company RQ score change between the years of 2017 to 2022?

    ```{reputation %>%}
      group_by(industry)
    ```

# Analysis plan

A plan for answering each of the questions including the variables involved, variables to be created (if any), external data to be merged in (if any).

1.  Variables involved: `name`, `industry`, `rating`
    1.  from the reputation dataset

    2.  create new average score variable for each rating attribute for each industry
2.  Variables involved: `company`, `rq`, `2022_rq`, `year`
    1.  from polls dataset

    2.  filtering for NA values, having to data wrangle for 2022 variable
3.  
