---
title: "Investigating Brand Reputation of America's Top 100 Companies"
subtitle: "Proposal"
format: html
editor: visual
---

```{r, include=FALSE}
#| label: load-pkgs
#| message: false

library(tidyverse)
```

# Dataset

A brief description of your dataset including its provenance, dimensions, etc. as well as the reason why you chose this dataset.

Make sure to load the data and use inline code for some of this information.

```{r}
# poll <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2022/2022-05-31/poll.csv') 
# reputation <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2022/2022-05-31/reputation.csv')
```

```{r-load-data, include=FALSE}
poll <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2022/2022-05-31/poll.csv') 
reputation <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2022/2022-05-31/reputation.csv')
```

```{r loading poll data}
# write_csv(poll, "/home/guest/project-01-team_six/polls.csv")
```

```{r loading reputation data}
# write_csv(reputation, "/home/guest/project-01-team_six/reputation.csv")
```

We've interacted with many of these companies before, yet our familiarity with these entities is limited solely to their name recognition. In an effort to expand our knowledge and gain a more comprehensive understanding of their respective positions, we aim to delve deeply into the available data from the 2022 Axios-Harris Poll and draw deeper insights. We're interested in investigating companies across different industries that are highly reputable, but may have experienced either a decline or rise in the public's eye. 

# Questions

1.  Which industry on average ranked the highest for each rating attribute in the year of 2022?

2.  How does company RQ score change between the years of 2017 to 2022?

# Analysis plan

A plan for answering each of the questions including the variables involved, variables to be created (if any), external data to be merged in (if any).

Our analysis of question 1 will come from the `reputation` dataset. To see which industry on average ranked the highest for each attribute in 2022, we will use the variables `name`, `industry`, and `rating`. We will have to wrangle the data before visualization by creating an average rating variable for each attribute for each industry. Also, we may re-categorize some of the industries or place some companies in different industries to have better categories for comparison in the visualization. For our visualization, we are going to segment it by the seven attributes and show which industry has the highest average score in each.

1.  Variables involved: `company`, `rq`, `2022_rq`, `year`
    1.  from polls dataset

    2.  filtering for NA values, having to data wrangle for 2022 variable
2.  
